---
name: "Stephen M"
surname: "Pederson"
position: "Post-Doctoral Bioinformatician"
address: |
  | Dame Roma Mitchell Cancer Research Laboratories
  | Adelaide Medical School
  | University of Adelaide
  |
phone: +61 4 1333 9618
email: "stephen.pederson@adelaide.edu.au"
www: "researchers.adelaide.edu.au/profile/stephen.pederson"
github: steveped
linkedin: "stephen-pederson-06ba4914"
date: "`r format(Sys.time(), '%B %Y')`"
output: 
  vitae::moderncv:
    keep_tex: true
    theme: classic
---

<!-- Links not required in YAML. Keep here for future reference -->
<!-- orcid: 0000-0001-8197-3303 -->
<!-- twitter: StevePederson8 -->

<!-- Line spacing from detailed_entries() is too much. Reset here -->
\renewcommand*{\cventry}[7][-3mm]{%
  \cvitem[#1]{#2}{%
    {\bfseries#3}%
    \ifthenelse{\equal{#4}{}}{}{, {\slshape#4}}%
    \ifthenelse{\equal{#5}{}}{}{, #5}%
    \ifthenelse{\equal{#6}{}}{}{, #6}%
    .\strut%
    \ifx&#7&%
    \else{\newline{}\begin{minipage}[t]{\linewidth}\small#7\end{minipage}}\fi}}




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r packages}
library(vitae)
library(tidyverse)
library(magrittr)
library(scholar)
library(lubridate)
```


```{r getPubs}
sp <- "ZbO2KUEAAAAJ"
pubsRaw <- get_publications(sp, flush = TRUE) %>%
  as_tibble() %>%
  mutate(
    across(
      .cols = c("title", "author", "journal"), 
      .fns = stringi::stri_trans_general,
      id = "latin-ascii"
    )
  )
```

```{r getIF}
j <- unique(pubsRaw$journal) %>%
  .[. != ""] %>%
  str_subset("xiv$", negate = TRUE) %>%
  str_subset("Patent", negate = TRUE)
imp <- get_impactfactor(j) %>%
  mutate(
    j = j,
    ImpactFactor = case_when(
      # All taken from https://academic-accelerator.com/Impact-Factor-IF
      j %in% c("Plutella", "The Journal of Open Source Software") ~ NA_real_,
      j == "Oncotarget" ~ 3.71,
      j == "Epigenomes" ~ 3.83,
      j == "The Journal of Immunology" ~ 4.34,
      j == "Clinical & Translational Immunology" ~ 9.18,
      j == "International Journal of Nephrology and Renovascular Disease" ~ 2.59,
      !is.na(ImpactFactor) ~ ImpactFactor
    )
  )
pubsRaw %<>%
  left_join(
    dplyr::select(imp, journal = j, ImpactFactor),
    by = "journal"
  )
```

```{r shorten_full_authors}
shorten_full_authors <- function(x){
  x <- iconv(x, "latin1", "ASCII", "n")
  x <- stringi::stri_trans_general(x, id = "latin-ascii")
  x <- str_split(x, pattern = ",")
  x <- lapply(x, str_replace_all, pattern = "[a-z]+ ", replacement = " ")
  x <- lapply(x, str_trim)
  x <- lapply(x, str_replace_all, pattern = "([A-Z]) ([A-Z]) (.+)", replacement = "\\1\\2 \\3")
  vapply(x, str_flatten, character(1), collapse = ", ")
}
```

```{r replace_short, results='hide'}
needsFull <- !str_detect(pubsRaw$author, "Pederson")
full <- get_complete_authors(sp, pubsRaw$pubid[needsFull], initials = FALSE)
pubsRaw$author[needsFull] <- shorten_full_authors(full)
pubsRaw$author <- str_replace_all(pubsRaw$author, "(SC B )(T Sadlon.+)", "\\2, SC Barry")
```


```{r pred_h}
pred_h <- predict_h_index(sp)
```

```{r finalPubs}
pubs <- pubsRaw %>%
  dplyr::filter(
    # cites > 1 | year == "2020", #year(today()),
    !str_detect(journal, "(Patent|EUROPEAN|REPRODUCTIVE|Wiley)")
  ) %>%
  mutate(
    author = str_replace_all(author, "\\.\\.\\.", "\\\\emph{et al}"),
    author = str_replace_all(author, " &", " \\&"),
    title = str_replace_all(title, " &", " \\&"),
    title = case_when(
      title == "Brain Transcriptome Analysis of a Protein-Truncating Mutation in Sortilin-Related Receptor 1 Associated With Early-Onset Familial Alzheimer's Disease Indicates Early Effects on ..." ~ str_replace(title, "\\.{3}", "Mitochondrial and Ribosome Function"),
      title == "Brain transcriptome analysis reveals subtle effects on mitochondrial function and iron homeostasis of mutations in the SORL1 gene implicated in early onset familial Alzheimer's ..." ~ str_replace(title, "\\.{3}", "Disease"),
      title == "Transcriptome analyses of 7-day-old zebrafish larvae possessing a familial Alzheimer's disease-like mutation in psen1 indicate effects on oxidative phosphorylation, mcm ..." ~ str_replace(title, "\\.{3}", "functions, and iron homeostasis"),
      TRUE ~ title
    ),
    journal = str_replace_all(journal, " &", " \\\\&"),
    journal = str_to_title(journal),
    journal = str_replace_all(journal, "^Bmc", "BMC"),
    journal = str_replace_all(journal, "^Plos", "PLOS"),
    journal = str_replace_all(journal, "^Mbio", "mBio"),
    journal = str_replace_all(journal, "Dna$", "DNA"),
    journal = str_replace_all(journal, "Biorxiv", "bioRxiv"),
    journal = str_replace_all(journal, "Medrxiv", "medRxiv"),
    journal = case_when(
      grepl("BMEA", title) ~ "PhD Thesis",
      !grepl("BMEA", title) ~ journal
    )
  ) 
wif <- weighted.mean(pubs$ImpactFactor, pubs$cites, na.rm = TRUE)
```



```{r pub2Tex}
pub2Tex <- function(x){
  paste(
    "\t",
    "\\item",
    str_replace_all(x$author, "SM* Pederson", "\\\\textbf{SM Pederson}") %>%
      str_replace_all("Pederson SM*", "\\\\textbf{Pederson SM}"),
      # str_replace_all("‐", "-"),
    paste0("(", x$year, ")"),
    paste0(
      "\\emph{", x$title, "}"
    ),
    paste0(
      "\\textbf{", x$journal, "}"
    ),
    x$number,
    # paste0(
    #   "(\\emph{",
    #   x$cites,
    #   " citations})"
    # ),
    "\\\\[-3mm]",
    "\n"
  ) 
}
```


# Profile

 * Whilst training in the context of the immune dysregulation of T cells, a keen interest of mine is *Transcriptomics and Gene Regulatory Networks*. As part of my PhD (2018), I developed a novel, Bayesian statistical methodology for the analysis of Whole Transcript Microarray data, focussing on the detection of alternate isoform usage. As a consequence, a potential role for ncRNA in the activation of T cells was revealed for the first time.
 * As the inaugural co-ordinator of the Bioinformatics Hub at the University of Adelaide (2014-2020), *I oversaw the growth of this facility from a sole employee in 2014 to 4.0FTE staff in core positions by 2020*, along with a further 12 contract and casual staff. 
 <!-- Collectively, our team were named investigators on $8.5m in research funding. -->
 * Bioinformatics education and training have always been an important interest. Locally *I have delivered bioinformatics training workshops to over 1500 staff and students*, as well as delivering training workshops at national training events. From the years 2016-2018, I ran the *largest training event in the language `R` held in Australia*, attracting attendees from around Australia. I have also run corporate R training for both Sugar Research Australia (QLD) and SARDI
 * I played an active role in the *establishment of an Undergraduate Major in Bioinformatics* (B.Sc), acting as Course Convenor for Transcriptomics Applications III in 2020, writing and delivering nearly all lecture and practical material. 
 * I am an enthusiastic supporter of open source software and am the senior author/developer of *two Bioconductor packages*. From 2016-2019, I played an active role in establishing the international event BiocAsia, bringing the training and networking opportunities of the Bioconductor community to our local region. 
 <!-- * Having published `r nrow(dplyr::filter(pubs, !str_detect(journal, "iv$")))` peer-reviewed papers, the average Impact Factor across all my accepted publications is currently `r round(wif, 2)` with over 500 citations. Despite being an ECR of 2 years, I have an *h-index* of `r pred_h$h_index[1]`. -->
 * Prior to bioinformatics, I was a professional musician and remain active in the local arts community, having performed at WOMADelaide four times

\small

# Employment History

```{r}
tribble(
  ~Start, ~End, ~Org, ~Role,
  1992, 2008, "", "Self-Employed Musician",
  2011, 2014, "School of Mathematical Sciences, University of Adelaide", "Casual Tutor",
  2014, 2020, "Bioinformatics Hub, University of Adelaide", "Co-ordinator",
  2020, NA_real_, "Dame Roma Mitchell Cancer Research Laboratories, Univerrsity of Adelaide", "Post-Doctoral Bioinformatician",
) %>% 
  arrange(desc(Start)) %>%
  mutate(
    Year = case_when(
      Start == End ~ as.character(Start),
      is.na(End) ~ paste0(Start,"-"),
      Start != End ~ paste(Start, End, sep = "-")
    ),
    Loc = "Adelaide, South Australia"
  ) %>%
  detailed_entries(
    Role, Year, Org, where = Loc
  )
```


# Education

```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Bachelor of Music", "1989-1991", "University of Adelaide", "South Australia, Australia",
  "Bachelor of Science", "2002-2005", "University of Adelaide", "South Australia, Australia",
  "Bachelor of Mathematics & Comp. Sc. (Hons)", "2006-2007", "University of Adelaide", "South Australia, Australia",
  "Doctor of Philosophy (Medicine)", "2008-2018", "University of Adelaide", "South Australia, Australia"
) %>% 
  arrange(desc(Year)) %>%
  detailed_entries(Degree, Year, Institution, Where)
```


# Presentations

```{r}
tribble(
  ~Year, ~Event, ~Location, ~Title, 
  2019, "Bioconductor", "New York, NY", "Short Talk & Poster - ngsReports and strandCheckR: Two new packages for quality control of NGS data",
  2019, "BiocAsia/BioinfoSummer", "Sydney", "Short Talk - Experiences of a First-Time Package Submitter",
  2017, "BioinfoSummer", "Adelaide", "Training Workshop - Data Visualisation in R",
  2016, "ABACBS", "Brisbane", "Poster - Use of a weighted bootstrap for indentifying the regulatory role of genomic repeat elements",
  2015, "NCBI", "Bethesda, MD", "Presentation - Bioinformatics from Down Under",
  2015, "BioinfoSummer", "Sydney", "Training Workshop - Introduction to R"
) %>%
  arrange(desc(Year)) %>%
  detailed_entries(
    Event,
    Year,
    Location,
    Title
  )
```


# Community Engagement

```{r}
tribble(
  ~Start, ~End, ~Org, ~Role,
  2018, 2021, "Australian Bioinformatics And Computational Biology Society", "Ordinary Member",
  2016, 2019, "Bioconductor", "BiocAsia Conference Committee",
  2017, 2017, "Australian Bioinformatics And Computational Biology Society", "Conference Committee",
  2016, 2016, "Australian Mathematical Sciences Institute", "BioinfoSummer Conference Committee",
  2014, 2016, "Australasian Genomics Technologies Association", "Ordinary Member",
  2014, 2015, "International Society for Computational Biology", "Ordinary Member"
) %>% 
  arrange(desc(End)) %>%
  mutate(
    Year = case_when(
      Start == End ~ as.character(Start),
      Start != End ~ paste(Start, End, sep = "-")
    )
  ) %>%
  brief_entries(
    Role,
    Year, 
    Org
  )
```

# Technical Skills

 * Highly fluent in the language `R` and using High-Performance Computing systems, along with `snakemake` for pipeline management
 * Experienced with `linux` (Ubuntu), `bash`, `Git` for version control, `markdown` and \LaTeX
 * A working familiarity with `python` and `C/C++`
 * Expertise in RNA-Seq, scRNA-Seq, ChIP-Seq, RAD/GBS-Seq, HiC-Seq and microarray data


# Publications Since 2018

\begin{enumerate}
  \small
```{r, results='asis'}
printPubs <- pubs %>%
  arrange(desc(year)) %>%
  dplyr::filter(year >= 2018) %>%
  split( f = seq_len(nrow(.))) %>%
  lapply(pub2Tex)
for (i in seq_along(printPubs)) cat(printPubs[[i]])
```

\end{enumerate}

```{r, eval=FALSE}
# Just doesn't look as nice as my manual version
bibliography_entries("scholar.bib") %>%
  arrange(issued, desc(author$family)) %>%
  dplyr::filter(
    !is.na(`container-title`),
    !str_detect(`container-title`, "EUROPEAN"),
    !type %in% c("no-type", "paper-conference")
  )
```

